---
title: "assignment-3"
author: "Ronnie"
date: "4/9/2020"
output: 
  github_document:
    toc: true
    toc_depth: 3
---

## Load data

```{r message=FALSE, warning=FALSE, paged.print=FALSE, include=FALSE}
library(tmap)
library(sf)
library(tidyverse)
library(spdep)
library(patchwork)
```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
resale <- readRDS(here::here("data/resale_with_geom.rds"))
set.seed(13)
resale_sample <- sample_n(resale, 5000)

planning_areas <- st_read(here::here("data/MP14_shp/MP14_PLNG_AREA_NO_SEA_PL.shp")) %>%
  filter(!(OBJECTID == 49 | OBJECTID == 18))

hex_grid <- planning_areas %>%
  st_make_grid(st_bbox(.), square = FALSE, cellsize = 1500) %>%
  st_sf() %>%
  mutate(hex_id = row_number())

resale_hex <- st_join(resale_sample, hex_grid) %>%
  st_set_geometry(NULL)

```

## Average floor area

```{r}
hex_grid <- planning_areas %>%
  st_make_grid(st_bbox(.), square = FALSE, cellsize = 1500) %>% 
  st_sf() %>% 
  mutate(hex_id = row_number())

resale_hex <- st_join(resale_sample, hex_grid) %>% 
  st_set_geometry(NULL)

floor_area_sqm <- resale_hex %>% 
  group_by(hex_id) %>% 
  summarise(average_area = mean(floor_area_sqm)) %>% 
  left_join(hex_grid, .) %>% 
  #replace_na(list(floor_area_sqm = 999))
  filter(average_area > 0)
```


```{r}
hex_sp <- as(floor_area_sqm, 'Spatial')
hex_neighbors <- readRDS(here::here("data/hex_neighbors_connect_floor_area_sqm.rds"))
plot(hex_neighbors, coordinates(hex_sp))
```

```{r}
cluster_data <- floor_area_sqm %>% 
  st_set_geometry(NULL) %>%
  select(average_area)

hex_edge_costs <- nbcosts(hex_neighbors, cluster_data)
hex_edge_weights <- nb2listw(hex_neighbors, hex_edge_costs, style="B")
hex_mstree <- mstree(hex_edge_weights)
```

```{r}
hex_skater_k6 <- skater(hex_mstree[,1:2], 
                        cluster_data, 
                        ncuts = 5,
                        crit = 10)

floor_area_sqm$cluster <- as.character(hex_skater_k6$groups)
plot_k6 <- tm_shape(floor_area_sqm) + tm_polygons(col = "cluster")
plot_k6
```

```{r warning=FALSE}
floor_area_sqm %>% group_by(cluster) %>% 
  summarise(cluster_mean_area = round(mean(average_area), 2)) %>% 
  ggplot(aes(fct_reorder(cluster, cluster_mean_area), 
             cluster_mean_area)) + 
  geom_histogram(fill="#AED6F1", stat = "identity")+
  geom_text(aes(label=cluster_mean_area), size=3) +
  coord_flip()
```

**Discussion:** There average floor area have less difference between each other ranging from 80 sqm to 115 sqm. No surprisingly, Central Singapore (cluster 6) has least areas for housing because of land scarcity. The flat type, as such around 80 sqm in average, only enable to provide limit spaces which meet basic living functions. However, in northeast Singapore (Cluster 1), near Changi, Tampines and Pasir Ris, it has the largest areas in average, with North Singapore near Sembangwang, Yishun (Cluster 2) ranks the second in proximaty to the Northeast.

```{r}
hex_skater_k10 <- skater(hex_mstree[,1:2], 
                        cluster_data, 
                        ncuts = 9,
                        crit = 10)

floor_area_sqm$cluster <- as.character(hex_skater_k10$groups)
plot_k10 <- tm_shape(floor_area_sqm) + tm_polygons(col = "cluster")


hex_skater_k8 <- skater(hex_mstree[,1:2], 
                        cluster_data, 
                        ncuts = 7,
                        crit = 10)

floor_area_sqm$cluster <- as.character(hex_skater_k8$groups)
plot_k8 <- tm_shape(floor_area_sqm) + tm_polygons(col = "cluster")

hex_skater_k12 <- skater(hex_mstree[,1:2], 
                        cluster_data, 
                        ncuts = 11,
                        crit = 10)

floor_area_sqm$cluster <- as.character(hex_skater_k12$groups)
plot_k12 <- tm_shape(floor_area_sqm) + tm_polygons(col = "cluster")


tmap_arrange(plot_k6, plot_k8, plot_k10, plot_k12)
```

**Discussion: ** With less clusters, the central area will be more homogeneous. Also, the places near woodlands and Yishun were combined into one cluster in top-left chart. However, as the number of cluster increasing, the regions can differentiate from each other. 

## Average price per square meter

```{r}
hex_grid <- planning_areas %>%
  st_make_grid(st_bbox(.), square = FALSE, cellsize = 1500) %>% 
  st_sf() %>% 
  mutate(hex_id = row_number())

resale_hex <- st_join(resale_sample, hex_grid) %>% 
  st_set_geometry(NULL)

price_per_sqm <- resale_hex %>% 
  mutate(price_per_sqm = resale_price / floor_area_sqm) %>% 
  group_by(hex_id) %>% 
  summarise(price_per_sqm = mean(price_per_sqm)) %>% 
  left_join(hex_grid, .) %>% 
  #replace_na(list(floor_area_sqm = 999))
  filter(price_per_sqm > 0)
```


```{r}
hex_sp <- as(price_per_sqm, 'Spatial')
hex_neighbors <- readRDS(here::here("data/hex_neighbors_connect_price_per_sqm.rds"))
plot(hex_neighbors, coordinates(hex_sp))
```

```{r}
cluster_data <- price_per_sqm %>% 
  st_set_geometry(NULL) %>%
  select(price_per_sqm)

hex_edge_costs <- nbcosts(hex_neighbors, cluster_data)
hex_edge_weights <- nb2listw(hex_neighbors, hex_edge_costs, style="B")
hex_mstree <- mstree(hex_edge_weights)
```

```{r}
hex_skater_k6 <- skater(hex_mstree[,1:2], 
                        cluster_data, 
                        ncuts = 5,
                        crit = 10)

price_per_sqm$cluster <- as.character(hex_skater_k6$groups)
plot_k6 <- tm_shape(price_per_sqm) + tm_polygons(col = "cluster")
plot_k6
```

```{r warning=FALSE}
price_per_sqm %>% group_by(cluster) %>% 
  summarise(cluster_mean_price = round(mean(price_per_sqm), 2)) %>% 
  ggplot(aes(fct_reorder(cluster, cluster_mean_price), 
             cluster_mean_price)) + 
  geom_histogram(fill="#AED6F1", stat = "identity")+
  geom_text(aes(label=cluster_mean_price), size=3) +
  coord_flip()
```

**Discussion:** Unsurprisingly, central Singapore (cluster 2) is the most expensive and has huge different with other parts of Singapore at around 6000 per sqm. Apart from central Singapore, the prices shows less difference in a range between 3500 and 4800. The regions closer to each other show more similarity in prices. East Singapore (cluster 6) and northeast Singapore (cluster 4) take third and second places respectively. However,  west(cluster 3) and northwest (cluster 1 & 5) parts illustrate a lower price, comparing to east regions, especially the place near Chua Chu Kang (cluster 5) has lowest price at around 3673 per sqm in average. 

```{r}
hex_skater_k10 <- skater(hex_mstree[,1:2], 
                        cluster_data, 
                        ncuts = 9,
                        )

price_per_sqm$cluster <- as.character(hex_skater_k10$groups)
plot_k10 <- tm_shape(price_per_sqm) + tm_polygons(col = "cluster")


hex_skater_k8 <- skater(hex_mstree[,1:2], 
                        cluster_data, 
                        ncuts = 7,
                        )

price_per_sqm$cluster <- as.character(hex_skater_k8$groups)
plot_k8 <- tm_shape(price_per_sqm) + tm_polygons(col = "cluster")

hex_skater_k12 <- skater(hex_mstree[,1:2], 
                        cluster_data, 
                        ncuts = 11,
                        )

price_per_sqm$cluster <- as.character(hex_skater_k12$groups)
plot_k12 <- tm_shape(price_per_sqm) + tm_polygons(col = "cluster")


tmap_arrange(plot_k6, plot_k8, plot_k10, plot_k12)
```

**Discussion:** If it creates 6 clusters, the central area will be more homogeneous. Once the number of cluster increasing to 8, the regions differentiate from one single cluster and split into three parts: cluster 2, 7 an 8 in top-right chart with 8 clusters. However, with the number of clusters increasing to 10 and 12, there wound be less differentiation. 
